Magnetic Phase Transitions in an Aperiodic Monotile

### 2023-06-12 paper/main.tex
Done

\section{Introduction}
The Ising model was one of the first statistical mechanics models to contain a phase transition, and remains one of the only exactly solvable examples. It consists of defining a variable $s_i = \pm 1$ at every site of a lattice as a classical approximation of the up-or-down spin of a fermion. On further analogy with fermionic spins, it allows neighboring spins to interact under the classical Hamiltonian
\begin{equation}
  H = -J \sum_{\braket{ij}} s_i s_j - h \sum_i s_i
  \label{eqn:ising}
\end{equation}
where $h$ is an applied magnetic field and $\braket{ij}$ indicate neighbors. This model's phase transition is visible in the magnetization, defined as
\begin{equation}
  M = \frac{1}{N}\Braket{\sum_{i} s_i}
  \label{eqn:magnetization}
\end{equation}
where the brackets indicate a statistical average over all states, weighted by the Boltzmann factor $e^{-\beta H}$.

@@

The phenomenon of critical temperature has been generalized to other theories. In particular, if the system's variables are arranged as a members of a field $s(\bm x)$, then the free energy may be approximated as an integral over functions of $s(\bm x)$. This approximation is called Landau Field Theory and it yields a form which is readily able to create phase transitions. Taking the Ising model as an example, we approximate the $s_i$ spins defined on discrete points $x_i$ as the field $s(\bm x)$ defined for any position $\bm x$. The original Hamiltonian contained translational, rotational, and $s\rightarrow-s$ symmetry, and therefore the free energy should as well. Expanding its integrand in powers of $s(\bm x)$ guarantees the form
\begin{equation}
F \propto -\int d^2 x \brackets{(\nabla s(\bm x))^2 + A s(\bm x)^2 + B s(\bm x)^4 + \dots}
\label{eqn:landau}
\end{equation}
where $A$ and $B$ are arbitrary constants. If $A>0$, then free energy is minimized at every point by setting $s(\bm x)=0$. However if $A<0$, then the average value of $s(\bm x)0$ is nonzero and the crystal is magnetized. In principle, all the parameters of the model including $A$ may depend on temperature, so any theory whose $A$ crosses from positive to negative contains a phase transition.

The utility of this field theory framework is in its generality. All Ising-like models are Landau field theories and are all described by Eq~\ref{eqn:landau}, meaning they have the same near-critical-point behavior. This statement is independent of the lattice or interaction strengths; it depends only on the field $\bm s$ and the dimensionality of the lattice. A particular lattice's only unique properties its its values of $A$ and $B$; i.e., the value of its critical temperature.

@@

In complicated crystals, the only available approaches to computing this $T_c$ are numerical. For example, the Penrose lattice consists of two tiles which fill the two-dimensional plane aperiodically, making it difficult to model the system analytically. Ref. \cite{penrose-ising} computes magnetism as a function of temperature using numerical methods and finds that the aperiodic nature of the lattice does not affect the physics at a qualitative level, since $M(T)$ inherits most of its properties from the general Landau theory.

@@

In this final project, we use numerical methods to compute the phase diagram $T_c(a/b)$ for the einstein tile under the Ising model and two similar models. The spin models are discussed in section \ref{sec:spin-models} and the numerical methods in section \ref{sec:wolff}. We also compute the phase diagram for a rectangular lattice to compare to the einstein lattice phase diagram in section \ref{sec:results}, later using this calculation to solve for the one-dimensional transverse Ising model in section \ref{sec:tim}. We conclude with section \ref{sec:conclusion}.

@@

In addition to the Ising model Hamiltonian (Eq.~\ref{eqn:ising}), we study the XY model which consists of promoting $s_i$ to a two-dimensional unit vector defined at every lattice point. The interaction term in the Hamiltonian becomes $\bm s_i \cdot \bm s_j$. A third spin model is the Heisenberg model, which is similar to the XY model except that $s_i$ is promoted to a three-dimensional unit vector. These models are all different in that they have progressively more spin degrees of freedom (DOFs), leading to progressively lower $T_c$. Each model also has specialized features, like the XY model's topologically protected vortices and the Heisenberg model's pseudo-phase transition \cite{tomita2014finite}, but these are beyond the scope of this report.

@@

Beyond the magnetization $M$, another interesting property of a spin model is its susceptibility $\chi = \frac{dM}{dh}$. Even in a model which sets $h=0$, $\chi$ can be achieved by explicitly differentiating Eq.~\ref{eqn:magnetization}, including the $e^{-\beta H}$ weight in the average. This differentiation results in the zero-$h$ limit of
\begin{equation}
  \chi = \frac{\partial M}{\partial h} = \frac{1}{N}\Braket{\sum_i(s_i - M)^2}.
\end{equation}
This is a special case of the fluctuation-dissipation theorem, which states that a driving field $h$ causes fluctuations in the driven variable $M$ which are proportional to the susceptibility $\chi$. A spin model is most susceptible to magnetic field near $T_c$, because the presence of a magnetic field changes the value of $T_c$ and therefore can greatly change the magnetization of a system near $T_c$. In particular, the susceptibility diverges at $T_c$ and otherwise falls off to zero.

@@

For an MCMC to be correct, $P(\phi \rightarrow \psi)$ must be reversible, meaning that the probability to go from $\phi$ to $\psi$ must be equal to the probability to go from $\psi$ to $\phi$. This condition is necessary to prevent $K$ from getting stuck in an unrepresentative region of phase space. The probability to go from $\phi$ to $\psi$ is the probability for the MCMC to be at state $\phi$ in the first place times the probability to switch to $\psi$. Thus, detailed balance requires
\begin{equation}
  P(\phi \rightarrow \psi)P(\phi) = P(\psi \rightarrow \phi)P(\psi)
  \label{eqn:detailed-balance}
\end{equation}
Any method of generating a new state $\psi$ from the old state $\phi$ that satisfies Eq.~\ref{eqn:detailed-balance} is a valid MCMC.

There are many methods of generating new states, but the one we will use in this paper is the Wolff algorithm, which is known to converge well near the critical temperature \cite{wolff1989collective}. The algorithm generates a new state $\psi$ by probabilistically choosing a cluster of spins and then flipping all the spins at once. The cluster starts the cluster at a random location $i$. It then adds each neighbor $j$ to the cluster with probability
\begin{equation}
  P(i,j) = 1 - \exp\parens{\mathrm{min}\left[0, 2\beta s_i s_j \right]}.
  \label{eqn:wolff-algo}
\end{equation}
For every spin $j$ which was added to the cluster, it then uses the same equation to assign $j$'s neighbors to the cluster until every neighbor has been rejected. At that point, the cluster's spins are flipped to form the new state $\psi$. Ref.~\cite{wolff1989collective} confirms that Eq.~\ref{eqn:wolff-algo} satisfies the detailed balance constraint and therefore is a good MCMC.

@@

We implemented the Wolff algorithm with a custom code base \footnote{\url{https://github.com/jack-dinsmore/quasing-model}} and executed for a variety of spin models and crystals. Except when stated otherwise, we used a finite-size lattice with initial conditions of random spins. Magnetization $M$ and susceptibility $\chi$ measurements were collected every 16 cluster flips to prevent autocorrelation. 10,000 $M$ and $\chi$ measurements were made when $M$ was small, and when $M$ was large we made 1,000 measurements because the algorithm was found to be more stable for $T< T_c$.

@@

\begin{figure*}
  \centering
  \includegraphics[width=0.49\linewidth]{../figs/square.pdf}\hfill
  \includegraphics[width=0.49\linewidth]{../figs/penrose.pdf}
  \caption{Magnetization and susceptibility as a function of $T_c$ under the XY and Ising models. \textit{Left}: A square lattice with $N=128^2=16,384$ sites. \textit{Right}: A Penrose lattice with $N=21,316$ sites. The three spin models yield very different predictions but the two crystals are similar due to their field theory explanation.}
  \label{fig:square}
\end{figure*}

@@

For the sake of approximating $T_c$, we take the $T$ point at which the steepest line tangent to the $M(T)$ curve intersects $M=0.5$. The critical temperatures obtained from this approximation are shown in table \ref{tab:tc}. These temperatures are very close to the true values known for the square lattice: For the Ising model, the Onsager solution gives $T_c = \frac{2}{\ln(1 + \sqrt{2})} \approx 2.2691$ in units where $J=k_B=1$, and for the XY model MCMC methods give $T_c \approx 0.8816$. The Heisenberg model does not have a true phase transition, but this fact is difficult to verify in a lattice of finite size, which never have true phase transitions in any case. 
\begin{table}
  \centering
  \begin{tabular}{c|ccc}
    \hline \hline
    & Square & Penrose & Einstein\\ \hline
    Ising & 2.26 & 2.36 & 0.396\\
    XY & 0.928 & 0.943 & 0.130\\
    Heisenberg & 0.493 & 0.462 & 0.0548\\ \hline \hline
  \end{tabular}
  \caption{Critical temperatures $T_c$ for the Ising, XY, and Heisenberg models defined on the square and Penrose lattices in units where $J = k_B = 1$. These values are very close to the true values (see text). Also shown are $T_c$ for the einstein lattice with $\eta=-\frac{1}{3}$ as depicted in figure \ref{fig:einstein}.}
  \label{tab:tc}
\end{table}

To demonstrate that the $M(T)$ relationship is not strongly dependent on the lattice, we also compute the same quantities for the Penrose lattice, shown in the right two panels of figure \ref{fig:square}. The same limiting behavior of $\chi$ and $M$ is present in both crystals, with slight changes in the steepness of the curves due to the difference in $N$. The major difference between the two is the actual value of $T_c$, which is slightly higher for Penrose (table \ref{tab:tc}). This confirms the result of Ref.~\cite{penrose-ising}.

@@

Figure \ref{fig:rect-phase} shows $T_c$ as a function of $\eta$ for the rectangular lattice under the Ising, XY, and Heisenberg models. As expected, the result is symmetric around $\eta \rightarrow -\eta$. Anisotropy appears to increase $T_c$ near $\eta = 0$ in the Ising and XY models, though this trend turns over at some $\eta$ to achieve the $T_c\rightarrow 0$ value in the limit as $\eta \rightarrow \pm 1$. For the Ising model, the turnover occurs at $\eta\approx \pm 0.625$ after which $T_c$ falls to zero rapidly. For the XY model, the maximum value of $T_c$ occurs at $\eta \approx \pm 0.362$, which is half the turnover $\eta$ of the Ising model. However, $T_c$ falls only slowly until $\eta \approx 0.5$, after which $T_c$ drops to zero.

@@

If $\eta$ is shifted by some amount $\eta \rightarrow \eta - x$, then $T_c(\eta)$ possesses $\eta\rightarrow -\eta$ symmetry. This is surprising because $\eta$ and $-\eta$ correspond to completely different lattices, suggesting the presence of some greater symmetry in the einstein tiling which has not yet been remarked. Other models defined on the einstein tiling do not contain this symmetry. The shift value $x$ is 0.029 for the Ising model, 0.052 for the XY model, and 0.033 for the Heisenberg model, though these shift values may depend on both numerical fluctuations of the MCMC and on the finite $N$ used.

@@

Quantum mechanical models such as the TIM cannot in general be represented by the same formalism as a classical model. However the the TIM is a special case: A $d$-dimensional TIM is equivalent to a $d+1$-dimensional Ising model, as first discovered in Ref.~\cite{schultz64two}. We will prove this in the case of $d=1$. To see this, consider the quantum partition function
\begin{equation}
  Z = \mathrm{Tr}\parens{e^{-\tau H}}.
  \label{eqn:quantum-partition}
\end{equation}
For clarity, the Hamiltonian for the TIM is
\begin{equation}
  H = -J \sum_{\braket{ij}}\bm \sigma_i \cdot \bm \sigma_j - h \sum_i\sigma_i^3
\end{equation}
where $\bm \sigma_i$ is the spin operator (a vector of Pauli matrices) for site $i$. Since $H$ is time-independent and therefore self-commuting, this is 
\begin{equation}
  \begin{split}
    Z =& \mathrm{Tr}\parens{e^{-\epsilon H}\dots e^{-\epsilon H}}\\
     =& \sum_{\psi_0\in\mathcal{H}}\bra{\psi_0}e^{-\epsilon H}\parens{\sum_{\psi_1\in\mathcal{H}}\ket{\psi_1}\bra{\psi_1}}\times\dots\\
     &\times e^{-i\epsilon H}\parens{\sum_{\psi_n\in\mathcal{H}}\ket{\psi_n}\bra{\psi_n}}\ket{\psi_0}
  \end{split}
\end{equation}
where $\ket{\psi_a}$ indicates a quantum state in the Hilbert space $\mathcal{H}$. In the second line we just inserted the identity many times as well as the definition of the trace. The number $n$ is set such that $\epsilon n = \beta$ to satisfy Eq.~\ref{eqn:quantum-partition}. Setting $\epsilon$ sufficiently low, we can expand the exponential to first order in $\epsilon$:
\begin{equation}
  \begin{split}
    Z =& \sum_{\{\psi\}}\bra{\psi_0}(1 - \epsilon H)\ket{\psi_1}\bra{\psi_1}\times\dots\\
    &\times (1 - \epsilon H)\ket{\psi_n}\braket{\psi_n|\psi_0}.
  \end{split}
  \label{eqn:partition-simplified}
\end{equation}
where we have collapsed all the sums into one. In an orthonormal basis where $\braket{\psi_a|\psi_b} = \delta_{ab}$, the only $\mathcal{O}(\epsilon^0)$ term in Eq.~\ref{eqn:partition-simplified} is for all $\ket{\psi_a}$ equal. So to order $\mathcal{O}(\epsilon)$,
\begin{equation}
  \begin{split}
    Z &= 1 - \epsilon\sum_{a\neq b}\braket{\psi_a|H|\psi_b}\\
    Z &= 1 - \epsilon\sum_{a\neq b}\Braket{\psi_a|-J\sum_{\braket{ij}}\bm \sigma_i \cdot \bm \sigma_j - h\sum_i\sigma_i^3|\psi_b}
  \end{split}
\end{equation}
As our basis, we may take the wavefunctions $\ket{s_1 s_2 \dots s_N}$ where $s_i$ indicates the sign of the spin of site $i$ and there are $N$ sites. Explicit calculation verifies that $\braket{s_is_j|\bm \sigma_i \cdot \bm \sigma_j|s_i's_j'}=\delta_{s_i,s_j}\delta_{s'_i,s'_j}$. We sum over the $s'$ spins and replace the $s$ Kronecker delta with $s_i s_j$, which is equal to $\delta_{s_i,s_j}$ shifting by a constant. Likewise, $\braket{s|\sigma^3|s'}=\delta_{s,s'}\sim ss'$. Expressing $\ket{\psi_a}=\ket{s_1 s_2 \dots s_N}$ and $\ket{\psi_b}=\ket{s'_1 s'_2 \dots s'_N}$, we get
\begin{equation}
  Z = 1 - \epsilon\sum_{\{s\},\{s'\}} \brackets{-J\sum_{\braket{ij}}\bm s_is_j - h\sum_is_i s'_i}.
\end{equation}
Since this expression is order $\epsilon$, we may write it as an exponential:
\begin{equation}
  \begin{split}
    Z &= \sum_{\{s\},\{s'\}} e^{-K}.\\
    K &= -\epsilon J\sum_{\braket{ij}}\bm s_is_j +\frac{1}{2}\ln(\epsilon h)\sum_is_i s'_i.
  \end{split}
\end{equation}

@@

Phase diagrams of the critical temperature as a function of the lattice anisotropy were shown; these are not predicted by Landau field theory and are therefore less theoretically constrained.  Anisotropy was seen to increase the critical temperature of the square lattice, but decrease the critical temperature of the einstein tiling. The rectangular lattice achieved maximum critical temperature at model-dependent ``turnover values'' of anisotropy, and the einstein tiling displayed unexpected symmetry after a model-dependent shift in the anisotropy. Understanding either of these features may require deeper investigation.

>>>>

\section{Introduction}
The Ising model was one of the first statistical mechanics models to contain a phase transition and remains one of the only exactly solvable examples. It consists of defining a variable $s_i = \pm 1$ at every site of a lattice as a classical approximation of the up-or-down spin of a fermion. On further analogy with fermionic spins, it allows neighboring spins to interact under the classical Hamiltonian
\begin{equation}
  H = -J \sum_{\braket{ij}} s_i s_j - h \sum_i s_i
  \label{eqn:ising}
\end{equation}
where $h$ is an applied magnetic field and $\braket{ij}$ indicate neighbors. Henceforth, we'll set $h=0$. This model's phase transition is visible in the magnetization, defined as
\begin{equation}
  M = \frac{1}{N}\Braket{\sum_{i} s_i}
  \label{eqn:magnetization}
\end{equation}
where the brackets indicate a statistical average over all states, weighted by the Boltzmann factor $e^{-\beta H}$.

@@

The phenomenon of critical temperature has been generalized to other theories. In particular, if the system's variables are arranged as a members of a field $s(\bm x)$, then the free energy may be approximated as an integral of a function of $s(\bm x)$. This approximation is called Landau Field Theory and it yields a form which is readily able to create phase transitions. Taking the Ising model as an example, we approximate the $s_i$ spins defined on discrete points $x_i$ as the field $s(\bm x)$ defined for any position $\bm x$. The original Hamiltonian contained translational, rotational, and $s\rightarrow-s$ symmetry, and therefore the free energy should as well. Expanding its integrand in powers of $s(\bm x)$ guarantees the form
\begin{equation}
F \propto -\int d^2 x \brackets{(\nabla s(\bm x))^2 + A s(\bm x)^2 + B s(\bm x)^4 + \dots}
\label{eqn:landau}
\end{equation}
where $A$ and $B$ are arbitrary constants. If $A>0$, then the integrand is minimized by setting $s(\bm x)=0$. However if $A<0$, then the average value of $s(\bm x)$ is nonzero and the crystal is magnetized. In principle, all the parameters of the model including $A$ may depend on temperature, so any theory whose $A$ crosses from positive to negative at a temperature $T_c$ contains a phase transition.

The utility of this field theory framework is its generality. All Ising-like models are Landau field theories and are all described by Eq~\ref{eqn:landau}, meaning they have the same near-critical-point behavior. This statement is independent of the lattice or interaction strengths; it depends only on the degrees of freedom (DOFs) of the field $s$ and the dimensionality of the lattice. A particular lattice's only unique properties its its values of $A$ and $B$; i.e., the value of its critical temperature.

@@

In complicated crystals, the only available approaches to computing $T_c$ are numerical. For example, the Penrose lattice consists of two tiles which fill the two-dimensional plane aperiodically, making it difficult to model the system analytically. Ref. \cite{penrose-ising} computes magnetism as a function of temperature using numerical methods and finds that the aperiodic nature of the lattice does not affect the physics at a qualitative level, since $M(T)$ inherits most of its properties from the general Landau theory.

@@

In this final project, we use numerical methods to compute the phase diagram $T_c(a/b)$ for the einstein tile under the Ising model and two similar models. The spin models are discussed in section \ref{sec:spin-models} and the numerical methods in section \ref{sec:wolff}. We also compute the phase diagram for a rectangular lattice to compare to the einstein lattice phase diagram in section \ref{sec:results}, later using this calculation to explain the one-dimensional transverse Ising model in section \ref{sec:tim}. We conclude with section \ref{sec:conclusion}.

@@

In addition to the Ising model Hamiltonian (Eq.~\ref{eqn:ising}), we study the XY model which consists of promoting $s_i$ to a two-dimensional unit vector defined at every lattice point. The interaction term in the Hamiltonian becomes $\bm s_i \cdot \bm s_j$. A third spin model is the Heisenberg model, which promotes $s_i$ to a three-dimensional unit vector. These models are all different in that they have progressively more spin DOFs, leading to progressively lower $T_c$. Each model also has specialized features, like the XY model's topologically protected vortices and the Heisenberg model's pseudo-phase transition \cite{tomita2014finite}, but these are beyond the scope of this report.

@@

Beyond the magnetization $M$, another interesting property of a spin model is its susceptibility $\chi = \frac{dM}{dh}$. Even in a model which sets $h=0$, $\chi$ can be achieved by explicitly differentiating Eq.~\ref{eqn:magnetization}, including the $e^{-\beta H}$ weight in the average. This differentiation results in the zero-$h$ limit of
\begin{equation}
  \chi = \frac{\partial M}{\partial h} = \frac{1}{N}\Braket{\sum_i(s_i - M)^2}.
\end{equation}
This is a special case of the fluctuation-dissipation theorem, which states that a driving field $h$ causes fluctuations in the driven variable $M$ which are proportional to the susceptibility $\chi$. A spin model is most susceptible to magnetic field near $T_c$, because the presence of a magnetic field changes the value of $T_c$ and therefore can greatly change the magnetization of a system near $T_c$. Thus the susceptibility diverges at $T_c$ and otherwise falls off to zero.

@@

For an MCMC to be correct, $P(\phi \rightarrow \psi)$ must be reversible, meaning that the probability to go from $\phi$ to $\psi$ must be equal to the probability to go from $\psi$ to $\phi$. This condition is necessary to prevent $K$ from getting stuck in an unrepresentative region of phase space. The probability to go from $\phi$ to $\psi$ is the probability for the MCMC to be at state $\phi$ in the first place times the probability to switch to $\psi$. Thus, reversibility requires
\begin{equation}
  P(\phi \rightarrow \psi)P(\phi) = P(\psi \rightarrow \phi)P(\psi)
  \label{eqn:detailed-balance}
\end{equation}

There are many methods of generating new states, but the one we will use in this paper is the Wolff algorithm, which is known to converge well near the critical temperature \cite{wolff1989collective}. The algorithm generates a new state $\psi$ by probabilistically choosing a cluster of spins and then flipping all the spins at once. The cluster starts the cluster at a random location $i$. It then adds each neighbor $j$ to the cluster with probability
\begin{equation}
  P(i,j) = 1 - \exp\parens{\mathrm{min}\left[0, 2\beta s_i s_j \right]}.
  \label{eqn:wolff-algo}
\end{equation}
For every spin $j$ which was added to the cluster, it then uses the same equation to assign $j$'s neighbors to the cluster until every neighbor has been rejected. At that point, the cluster's spins are flipped to form the new state $\psi$. Ref.~\cite{wolff1989collective} confirms that Eq.~\ref{eqn:wolff-algo} satisfies the reversibility constraint Eq.~\ref{eqn:detailed-balance} and therefore is a good MCMC.

@@

We implemented the Wolff algorithm with a custom code base\footnote{\url{https://github.com/jack-dinsmore/quasing-model}} and executed it for a variety of spin models and crystals. Except when stated otherwise, we used a finite-size lattice with initial conditions of random spins. Magnetization $M$ and susceptibility $\chi$ measurements were collected every 16 cluster flips to prevent autocorrelation. 10,000 $M$ and $\chi$ measurements were made when $M$ was small, and when $M$ was large we made 1,000 measurements because the algorithm was found to be more stable for $T< T_c$.

@@

\begin{figure*}
  \centering
  \includegraphics[width=0.49\linewidth]{../figs/square.pdf}\hfill
  \includegraphics[width=0.49\linewidth]{../figs/penrose.pdf}
  \caption{Magnetization and susceptibility as a function of $T_c$ under the Ising, XY, and Heisenberg models. \textit{Left}: A square lattice with $N=128^2=16,384$ sites. \textit{Right}: A Penrose lattice with $N=21,316$ sites. The three spin models yield very different predictions but the two crystals are similar due to their field theory explanation.}
  \label{fig:square}
\end{figure*}

@@

For the sake of approximating $T_c$, we take the $T$ point at which the steepest line tangent to the $M(T)$ curve intersects $M=0.5$. The critical temperatures obtained from this approximation are shown in table \ref{tab:tc}. These temperatures are very close to the true values known for the square lattice: For the Ising model, the Onsager solution gives $T_c = \frac{2}{\ln(1 + \sqrt{2})} \approx 2.2691$ in units where $J=k_B=1$, and for the XY model larger lattice MCMC methods give $T_c \approx 0.8816$. The $N\rightarrow \infty$ Heisenberg model does not have a true phase transition, but this fact is difficult to verify in a lattice of finite size, which never have true phase transitions in any case. 
\begin{table}
  \centering
  \begin{tabular}{c|ccc}
    \hline \hline
    & Square & Penrose & Einstein\\ \hline
    Ising & 2.26 & 2.36 & 0.396\\
    XY & 0.928 & 0.943 & 0.130\\
    Heisenberg & 0.493 & 0.462 & 0.0548\\ \hline \hline
  \end{tabular}
  \caption{Critical temperatures $T_c$ for the Ising, XY, and Heisenberg models defined on the square and Penrose lattices in units where $J = k_B = 1$. These values are very close to the true values (see text). Also shown are $T_c$ for the einstein lattice with $\eta=-\frac{1}{3}$ as depicted in figure \ref{fig:einstein}.}
  \label{tab:tc}
\end{table}

To demonstrate that the $M(T)$ relationship is not strongly dependent on the lattice, we also compute the same quantities for the Penrose lattice, shown in the right two panels of figure \ref{fig:square}. The same limiting behavior of $\chi$ and $M$ is present in both crystals, with slight changes in the steepness of the curves due to the difference in $N$. The major difference between the two is the actual value of $T_c$, which is slightly higher for Penrose (table \ref{tab:tc}).

@@

Figure \ref{fig:rect-phase} shows $T_c$ as a function of $\eta$ for the rectangular lattice under the Ising, XY, and Heisenberg models. As expected, the result is symmetric around $\eta \rightarrow -\eta$. Anisotropy increases $T_c$ near $\eta = 0$ in the Ising and XY models, though this trend turns over at some $\eta$ to achieve $T_c\rightarrow 0$ in the limit as $\eta \rightarrow \pm 1$. For the Ising model, the turnover occurs at $\eta\approx \pm 0.625$ after which $T_c$ falls to zero rapidly. For the XY model, the maximum value of $T_c$ occurs at $\eta \approx \pm 0.362$, which is half the turnover $\eta$ of the Ising model. However, $T_c$ falls only slowly until $\eta \approx 0.5$, after which $T_c$ drops to zero.

@@

If $\eta$ is shifted by some amount $\eta \rightarrow \eta - x$, then $T_c(\eta)$ possesses $\eta\rightarrow -\eta$ symmetry for the einstein tiling. This is surprising because $\eta$ and $-\eta$ correspond to completely different lattices, suggesting the presence of some greater symmetry in the einstein tiling which has not yet been remarked. Other models defined on the einstein tiling do not always contain this symmetry. The shift value $x$ is 0.029 for the Ising model, 0.052 for the XY model, and 0.033 for the Heisenberg model, though these shift values may depend on both numerical fluctuations of the MCMC and on the finite $N$ used.

@@

Quantum mechanical models such as the TIM cannot in general be represented by the same formalism as a classical model. However the the TIM is a special case: A $d$-dimensional TIM is equivalent to a $d+1$-dimensional anisotropic Ising model, as first discovered in Ref.~\cite{schultz64two}. We will prove this in the case of $d=1$. To see this, consider the quantum partition function
\begin{equation}
  Z = \mathrm{Tr}\parens{e^{-\beta H}}.
  \label{eqn:quantum-partition}
\end{equation}
For clarity, the Hamiltonian for the TIM is
\begin{equation}
  H = -J \sum_{\braket{ij}}\bm \sigma_i \cdot \bm \sigma_j - h \sum_i\sigma_i^3
\end{equation}
where $\bm \sigma_i$ is the spin operator (a vector of Pauli matrices) for site $i$. Since $H$ is time-independent and therefore self-commuting, this is 
\begin{equation}
  \begin{split}
    Z =& \mathrm{Tr}\parens{e^{-\epsilon H}\dots e^{-\epsilon H}}\\
     =& \sum_{\psi_0\in\mathcal{H}}\bra{\psi_0}e^{-\epsilon H}\parens{\sum_{\psi_1\in\mathcal{H}}\ket{\psi_1}\bra{\psi_1}}\times\dots\\
     &\times e^{-i\epsilon H}\parens{\sum_{\psi_n\in\mathcal{H}}\ket{\psi_n}\bra{\psi_n}}\ket{\psi_0}
  \end{split}
\end{equation}
where $\ket{\psi_a}$ indicates a quantum state in the Hilbert space $\mathcal{H}$. In the second line we just inserted the identity many times as well as the definition of the trace. The number $n$ is set such that $\epsilon n = \beta$ to satisfy Eq.~\ref{eqn:quantum-partition}. Setting $\epsilon$ sufficiently low, we can expand the exponential to first order in $\epsilon$:
\begin{equation}
  \begin{split}
    Z =& \sum_{\{\psi\}}\bra{\psi_0}(1 - \epsilon H)\ket{\psi_1}\bra{\psi_1}\times\dots\\
    &\times (1 - \epsilon H)\ket{\psi_n}\braket{\psi_n|\psi_0}.
  \end{split}
  \label{eqn:partition-simplified}
\end{equation}
where we have collapsed all the sums into one. In an orthonormal basis where $\braket{\psi_a|\psi_b} = \delta_{ab}$, the only $\mathcal{O}(\epsilon^0)$ term in Eq.~\ref{eqn:partition-simplified} is for all $\ket{\psi_a}$ equal. So to order $\mathcal{O}(\epsilon)$,
\begin{equation}
  \begin{split}
    Z &= 1 - \epsilon\sum_{a\neq b}\braket{\psi_a|H|\psi_b}\\
    Z &= 1 - \epsilon\sum_{a\neq b}\Braket{\psi_a|-J\sum_{\braket{ij}}\bm \sigma_i \cdot \bm \sigma_j - h\sum_i\sigma_i^3|\psi_b}
  \end{split}
\end{equation}
As our basis, we may take the wavefunctions $\ket{s_1 s_2 \dots s_N}$ where $s_i$ indicates the sign of the spin of site $i$ and there are $N$ sites. Explicit calculation verifies that $\braket{s_is_j|\bm \sigma_i \cdot \bm \sigma_j|s_i's_j'}=\delta_{s_i,s_j}\delta_{s'_i,s'_j}$. We sum over the $s'$ spins and replace the $s$ Kronecker delta with $s_i s_j$, which is equal to $\delta_{s_i,s_j}$ shifted by a constant. Likewise, $\braket{s|\sigma^3|s'}=\delta_{s,s'}\sim ss'$. Expressing $\ket{\psi_a}=\ket{s_1 s_2 \dots s_N}$ and $\ket{\psi_b}=\ket{s'_1 s'_2 \dots s'_N}$, we get
\begin{equation}
  Z = 1 - \epsilon\sum_{\{s\},\{s'\}} \brackets{-J\sum_{\braket{ij}} s_is_j - h\sum_is_i s'_i}.
\end{equation}
Since this expression is order $\epsilon$, we may write it as an exponential:
\begin{equation}
  \begin{split}
    Z &= \sum_{\{s\},\{s'\}} e^{-K}.\\
    K &= -\epsilon J\sum_{\braket{ij}}\bm s_is_j +\frac{1}{2}\ln(\epsilon h)\sum_is_i s'_i.
  \end{split}
\end{equation}

@@

Phase diagrams of the critical temperature as a function of the lattice anisotropy were shown; these are not predicted by Landau field theory and are therefore less theoretically constrained.  Anisotropy was seen to increase the critical temperature of the square lattice but decrease the critical temperature of the einstein tiling. The rectangular lattice achieved maximum critical temperature at model-dependent ``turnover values'' of anisotropy, and the einstein tiling displayed unexpected symmetry after a model-dependent shift in the anisotropy. Understanding either of these features may require deeper investigation.
